---
title: "STOR565 Final Project EDA (PCA)"
output:
  html_document:
  latex_engine: xelatex
  pdf_document: default
date: "2023-03-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Testing

```{r}
library(magick)
library(jpeg)
# no.tumor1 = jpeg::readJPEG("C:/Users/Oscar/Downloads/BrainTumorClass/Training/no_tumor/1.jpg")
# no.tumor2 = jpeg::readJPEG("C:/Users/Oscar/Downloads/BrainTumorClass/Training/no_tumor/2.jpg")
# no.tumor3= jpeg::readJPEG("C:/Users/Oscar/Downloads/BrainTumorClass/Training/no_tumor/image (11).jpg")

# no.tumor1 = image_read("C:/Users/Oscar/Downloads/BrainTumorClass/Training/no_tumor/1.jpg")
# no.tumor1 = image_scale(no.tumor1,"512x512!")
# setwd("C:/Users/Oscar/Downloads/")
# image_write(no.tumor1, path = "notumor.jpg", format = "jpg")
```

## Extract Images

```{r}
#List of no_tumor files
path <- "C:/Users/Oscar/Downloads/BrainTumorClass/Training/no_tumor/"
files <- list.files(path=path, pattern="*.jpg") 
no.tumor <- list(rep(0,length(files)))

for(i in 1:length(files)){
  no.tumor[[i]] = readJPEG(paste0(path,files[i]))[,,1]
}

#glioma_tumor
path <- "C:/Users/Oscar/Downloads/BrainTumorClass/Training/glioma_tumor/"
files <- list.files(path=path, pattern="*.jpg") 
glioma.tumor <- list(rep(0,length(files)))

for(i in 1:length(files)){
  glioma.tumor[[i]] = readJPEG(paste0(path,files[i]))[,,1]
}

#meningioma_tumor
path <- "C:/Users/Oscar/Downloads/BrainTumorClass/Training/meningioma_tumor/"
files <- list.files(path=path, pattern="*.jpg") 
meningioma.tumor <- list(rep(0,length(files)))

for(i in 1:length(files)){
  meningioma.tumor[[i]] = readJPEG(paste0(path,files[i]))[,,1]
}

#pituitary_tumor
path <- "C:/Users/Oscar/Downloads/BrainTumorClass/Training/pituitary_tumor/"
files <- list.files(path=path, pattern="*.jpg") 
pituitary.tumor <- list(rep(0,length(files)))

for(i in 1:length(files)){
  pituitary.tumor[[i]] = readJPEG(paste0(path,files[i]))[,,1]
}
```

## Consider image dimensions

```{r}
tumor = c(no.tumor,glioma.tumor,meningioma.tumor,pituitary.tumor)
rows = rep(NA,length(tumor))
for(i in 1:length(tumor)){
  rows[i]=nrow(tumor[[i]])
}

cols = rep(NA,length(tumor))
for(i in 1:length(tumor)){
  cols[i]=ncol(tumor[[i]])
}

# Most images are square
median(rows/cols)
hist(rows/cols)

# Underestimated size of data
# len = round(median(rows))
# wid = round(median(cols))
```

## Rescale (Simplify) Images

```{r, warning=FALSE}
library(magick)

#List of no_tumor files
path <- "C:/Users/Oscar/Downloads/BrainTumorClass/Training/no_tumor/"
files <- list.files(path=path, pattern="*.jpg") 
setwd("C:/Users/Oscar/Downloads/RBrainTumorClass/Training/no_tumor/")

for(i in files){
  brain = image_read(paste0(path,i))
  brain = image_scale(brain,"100x100!")
  image_write(brain, path = i, format = "jpg")
}

#List of glioma_tumor files
path <- "C:/Users/Oscar/Downloads/BrainTumorClass/Training/glioma_tumor/"
files <- list.files(path=path, pattern="*.jpg") 
setwd("C:/Users/Oscar/Downloads/RBrainTumorClass/Training/glioma_tumor/")

for(i in files){
  brain = image_read(paste0(path,i))
  brain = image_scale(brain,"100x100!")
  image_write(brain, path = i, format = "jpg")
}

#List of meningioma_tumor files
path <- "C:/Users/Oscar/Downloads/BrainTumorClass/Training/meningioma_tumor/"
files <- list.files(path=path, pattern="*.jpg") 
setwd("C:/Users/Oscar/Downloads/RBrainTumorClass/Training/meningioma_tumor/")

for(i in files){
  brain = image_read(paste0(path,i))
  brain = image_scale(brain,"100x100!")
  image_write(brain, path = i, format = "jpg")
}

#List of pituitary_tumor files
path <- "C:/Users/Oscar/Downloads/BrainTumorClass/Training/pituitary_tumor/"
files <- list.files(path=path, pattern="*.jpg") 
setwd("C:/Users/Oscar/Downloads/RBrainTumorClass/Training/pituitary_tumor/")

for(i in files){
  brain = image_read(paste0(path,i))
  brain = image_scale(brain,"100x100!")
  image_write(brain, path = i, format = "jpg")
}
```

## New Extracted Images

```{r}
rm(list = ls())
library(jpeg)
#List of no_tumor files
path <- "C:/Users/Oscar/Downloads/RBrainTumorClass/Training/no_tumor/"
files <- list.files(path=path, pattern=".jpg") 
no.tumor <- list(rep(0,length(files)))

for(i in 1:length(files)){
  no.tumor[[i]] = readJPEG(paste0(path,files[i]))
}

#glioma_tumor
path <- "C:/Users/Oscar/Downloads/RBrainTumorClass/Training/glioma_tumor/"
files <- list.files(path=path, pattern=".jpg") 
glioma.tumor <- list(rep(0,length(files)))

for(i in 1:length(files)){
  glioma.tumor[[i]] = readJPEG(paste0(path,files[i]))
}

#meningioma_tumor
path <- "C:/Users/Oscar/Downloads/RBrainTumorClass/Training/meningioma_tumor/"
files <- list.files(path=path, pattern=".jpg") 
meningioma.tumor <- list(rep(0,length(files)))

for(i in 1:length(files)){
  meningioma.tumor[[i]] = readJPEG(paste0(path,files[i]))
}

#pituitary_tumor
path <- "C:/Users/Oscar/Downloads/RBrainTumorClass/Training/pituitary_tumor/"
files <- list.files(path=path, pattern=".jpg") 
pituitary.tumor <- list(rep(0,length(files)))

for(i in 1:length(files)){
  pituitary.tumor[[i]] = readJPEG(paste0(path,files[i]))
}

```

## PCA

```{r}
tumor = c(no.tumor,glioma.tumor,meningioma.tumor,pituitary.tumor)
brains = data.frame(tumor = c(rep("no_tumor",length(no.tumor)),
                              rep("glioma_tumor",length(glioma.tumor)),
                              rep("meningioma_tumor",
                                  length(meningioma.tumor)),
                              rep("pituitary_tumor",
                                  length(pituitary.tumor))))

pixel = matrix(data = rep(NA, length(tumor)*100*100),
                nrow = length(tumor),
                ncol = 100*100)
for(i in 1:100){
  for(j in 1:100){
    for(k in 1:length(tumor)){
      which.col = 100*(i-1)+j
      pixel[k,which.col] = tumor[[k]][i,j]
      
    }
  }
}
```

```{r,cache=TRUE}
pca.brain = prcomp(pixel, scale = FALSE)
var_explained_brain = data.frame(PC= paste0("PC",1:length(tumor)),                              var_explained=(pca.brain$sdev)^2/sum((pca.brain$sdev)^2),                   cum_explained=cumsum((pca.brain$sdev)^2)/sum((pca.brain$sdev)^2))
```

```{r}
lastPC = min(which(var_explained_brain$cum_explained>.9))

library(tidyverse)
var_explained_brain[1:9,] %>%
  ggplot(aes(x=PC,y=cum_explained, group=1))+
  geom_point(size=4)+
  geom_line()+
  labs(title="Scree plot: cum explained")
```

```{r, warning=FALSE}
PCN = c("tumor")
for(i in 1:lastPC){
  PC = pca.brain$x[,i]
  brains = cbind(brains,PC)
  PCN = append(PCN,paste0("PC",i))
}
colnames(brains)=PCN

#pca plot
library(ggplot2)
ggplot(brains, aes(x=PC1, y=PC2, color=tumor)) + 
    geom_point(size=2) 
setwd("C:/Users/Oscar/Downloads/")
```
## Rotation Matrix

PCA is computationally expensive, but necessary for dimensional reduction. However, if the testing set is used to calculate the rotation matrix, it may skew the error rate. Thus, it is important to consider how the rotation matrix may vary. Considering the variation is minimal, I will project onto the full training set. 

```{r}
pixel.train <- sample(2871,1000)
pixelII <- pixel[pixel.train,]
pca.brainII = prcomp(pixelII, scale = FALSE)

pca.diff <- pca.brain$rotation[,1:lastPC]-pca.brainII$rotation[,1:lastPC]
mean(pca.diff)
```

## KNN Classification Rate

Determine optimal k, approximate with 5-fold cross-validation. Accuracy decreases as k increases, at a decreasing rate. Theory: since classification is based on closest image, KNN is heavily dependent upon similar images. This indicates the model is not robust to outliers, possible evidence of overfitting.
```{r}
library(class)
knn.fcv <-function(df,k){
  class.accuracy = rep(NA,5)
  set = sample(rep(1:5,length.out=nrow(df)))
  for(i in 1:5){
    train = df[ifelse(set==i,F,T),]
    test = df[ifelse(set==i,T,F),]
    knn.pred = knn(train[,-1], test[,-1], train[,1], k)
    class.accuracy[i] = mean(knn.pred == test[,1])
  }
  return(mean(class.accuracy))
}

possible.k = seq(1,101,2)
accuracy.k = rep(NA,51)

for(i in 1:51){
  accuracy.k[i] = knn.fcv(brains,possible.k[i])
}
which.max(accuracy.k[1:10])
accuracy.k[1:10]
```

```{r}
knn.loocv = knn.cv(brains[,-1],brains[,1],which.max(accuracy.k))
mean(knn.loocv==brains[,1])
mean(knn.loocv[1:396]==brains[1:396,1])
mean(knn.loocv[397:1222]==brains[397:1222,1])
mean(knn.loocv[1222:2044]==brains[1222:2044,1])
mean(knn.loocv[2045:2871]==brains[2045:2871,1])
```